Estimating contemporary migration numbers of adults for iteroparous species based on kinship relationships found between populations

### 2022-03-09 latex/ms.tex
finish...

When both HS and PO pairs, which are found between populations, are available simultaneously, we can obtain efficient estimators by combined $\widehat{M_1}$ and $\widehat{M_2}$ for migration number: 
\begin{align}
\widehat{M_3} &= \frac{ N_{1} N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) },
\label{AP-1}
\end{align}
and by combining $\widehat{m_1}$ and $\widehat{m_2}$ for the migration rate: 
\begin{align}
\widehat{m_3} &= \frac{ N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right)}.
\label{AP-2}
\end{align}
For those estimators, $\widetilde{H}_{\rm HS,bet}$ and $\widetilde{H}_{\rm PO,bet}$ are weighted by sample size $n_{\rm O, 2}$ and $n_{\rm P, 2}$, respectively.

@@

We developed an individual-based model that tracks kinship relationships to evaluate the performance of the estimators. Here, the population structure was assumed to be identical to that in the development of the estimators. Population 1 and 2 comprised $N_1$ and $N_2$ parents with an equal sex ratio, and their offspring number was assumed to follow the geometric distribution with mean ${\bar \lambda_1}$ and ${\bar \lambda_2}$ (i.e., Poisson reproduction with mean $\lambda_{i,1}$ and $\lambda_{j,2}$, which follows the exponential distribution with mean ${\bar \lambda_1}$ and ${\bar \lambda_2}$), respectively. Migrant parents were selected randomly from population 1 at the end of the first year, and each offspring retained IDs of the parents, thereby making it possible to trace an HS and PO relationship.

@@

First, we evaluated the accuracy of ${\hat M}$s based on the relative bias. As expected, for most of the investigated parameter sets, we observed that their relative bias was less than 5\%, as shown in {\bf Table S1} in {\bf Supporting Information}. Thus, when the assumption that kinships are detected without any error is satisfied, it is reasonable to consider ${\hat M}$s nearly unbiased estimators.

@@

As the sample size or number of migrants increase, precision increases, and the shape of the distribution asymptotically becomes symmetric ({\bf Fig.~\ref{violins}}) because an increasing sample size or increasing number of migrants is likely to increase the observed number of kinship pairs found between the two populations ($\widetilde{H}_{\rm HS,bet}$ or $\widetilde{H}_{\rm PO,bet}$) and decreases the variance of those kinship pair numbers. In addition, the number of PO pairs in a given population ($\widetilde{H}_{\rm PO,1}$ or $\widetilde{H}_{\rm PO,2}$) contributes to the precision of estimating $N_1$ or $N_1$, respectively, thereby providing ${\hat M_4}$ and ${\hat M_5}$ with higher precision.

Note that our simulation can handle case with and without invasive sampling. Invasive sampling potentially affects the level of $m$ because the sampled parents from population 1 have no chance to move to population 2; thus, such parents cannot be sampled. In this case, $m$ should be defined by $M/(N_1-n_{P,1})$ rather than $M/N_1$, although this does not affect the estimator ${\hat M}$s, as demonstrated in {\bf Table S1} ({\bf Supporting Information}).

@@

Our primary contributions are summarized as follows. First, we formulated the probabilities of kinship pairs randomly selected between two populations ($\pi_{\rm HS,bet}$ and $\pi_{\rm PO,bet}$), which provides migration number or rate estimators. While similar derivations may be found in the (nongenetic) MR method, these are limited to cases where sampling is non-invasive and adult individuals are sampled. The proposed method can avoid these limitations, where such an advantages is characterized by the CKMR method for estimating population sizes. Second, we have demonstrated that the probabilities of kinship pairs are approximately independent of the reproductive potential; thus, information about it is not required to realize the estimation provided that migration is assumed to occur independently of the reproductive potential. This is a useful property of the estimators because it is natural that heterogeneity of reproductive potential within and between populations exits in iteroparous species but its observation is generally difficult. Third, we have demonstrated that estimators can be obtained using only genetic data (i.e., ${\hat M_5}$ and ${\hat m_4}$). There are several advantages in using the proposed estimator rather than estimating $\pi_{\rm HS,bet}$ (and/or $\pi_{\rm PO,bet}$) separately (via a genetic method) and $N_1$ (and/or $N_2$) (via a nongenetic method), including simplified sampling processes and analyzing designs and the availability of a unified framework of genetic analyses for detecting HS and PO pairs (similar discussion is found for estimating the ratio of effective breeding size to the census size, $N_{\rm b}/N$, in \cite{Akita_2020}).

@@

We believe there are several scenarios where the proposed estimator may be beneficial. The first case is where there is a large number of migrants between populations, e.g., $M>100$, which eliminates population differentiation and thus hampers the detection of the $M$ via population genetics method if $M$ is interpreted as an effective migration number. Information about movement between populations is essential to assess population dynamics in the conservation and management contexts, even if the migrant number is so large that it cannot be genetically assigned to two populations. The second case involves genetic monitoring conducted for each reproductive period (e.g., annually) because the proposed estimators consider information about contemporary migration that is explicitly specified the timing (e.g., year); thus, the time-series data for migration may reflect environmental changes, and this would provide insights into the underlying ecological processes. In addition, HS pairs found within the genetic population provide $N_{\rm b}$ \cite[]{wang2009new,WAPLES_2011}, which is also used to assess genetic health. The third is the case involves the development of integrated models that combine several data sources into a single analysis \cite[]{MAUNDER201361}. An example of this direction is to provide additional information about migration to ongoing CKMR projects for stock assessment in fisheries \cite[e.g.,][]{Bravington_2016,Hillary_2018,https://doi.org/10.1002/ece3.6296,ecolevol2021p,10.1093/icesjms/fsac002}, which would utilize kinship pairs for simultaneous estimation of population parameters. If the population structure is defined hypothetically, the data accumulated in such projects could be readily connected to our theory via likelihood methods. The fourth case is where sampling adults is difficult due to conservation practices or other reasons but instead sampling offspring is relatively easy. In such cases, although $N_2$ (and $N_1$) must be given externally to employ $\hat m_1$ (and $\hat M_1$), there is presently no existing method to estimate adult movement, especially when two populations cannot be genetically distinct, which indicates that the proposed method has the potential to expand the scope of current population monitoring techniques.

Finally, we discuss some caveats in terms of applying the proposed method. Our theory for developing the estimators assumes that kinships are detected without error. Many algorithms can be used to detect kinship pairs from single nucleotide polymorphisms or short tandem repeats \cite[e.g., ][]{Wang_2009Genetics, Huisman_2017}, although a HS pair requires many more DNA markers than a PO pair. In addition, iteroparous species may has several kinship types e.g., half-uncle--nephew or half-cousins, which are expected to appear frequently and should be accurately differentiated from HS pairs. It is desirable to estimate in advance how many markers are required for kinship detection in the target populations associated with simulation of pedigree reconstruction (e.g., \cite{Anderson_2020}). Note that the proposed estimators are limited to detecting parental movements on a period between breeding seasons in the given populations. Thus, the estimation of fine-scale spatiotemporal movements, which is available to integrate data sources \cite[]{Thorson_2021}, is beyond the scope of this paper. Although population dynamics models with coarser spatial resolution than the spatial scale of environmental layers are frequently used in assessment models, i.e., the target application of the proposed method, the proposed estimators require pre-specification of the population structure. Kinship relationships with sample location information potentially realize the ability to explore a plausible population structure by estimating the migration number or rate between hypothetical populations, which is expected to contribute the determination of a management unit. In addition, the current theory behind the estimators does not assume desynchronized reproduction within a population (e.g., skip spawning), correlation between individual mobility and fertility, and nonrandom sampling of parents (e.g., mothers with a large number of offspring are likely to be sampled). These issues will be the focus of future work.

@@

In the main text, we ignore the variation of reproductive potential among parents (i.e., both $\lambda_{i,1}$ and $\lambda_{j,2}$ are constant) to derive $\pi_{\rm HS,bet}$ and $\pi_{\rm PO,bet}$. Here, let $k_{i,1}$ and $k_{j,2}$ be the number of surviving offspring of parents $i$ and $j$ at sampling time in populations 1 and 2, and assumed to follow a kind of discrete distribution (e.g., Poisson or negative binomial distribution) with mean $\lambda_{i,1}$ and $\lambda_{j,2}$, respectively. Without loss of generality, we set the index such that parents with $i=1$ to $M$ in population 1 and $j=1$ to $M$ in population 2 are identical migrants. For example, the parent with $i=1$ reproduces $k_{1,1}$ offspring (in population 1) and then reproduce $k_{1,2}$ offspring (in population 2) after migration. Given $k_{i,1}$ and $k_{j,2}$, the conditional probability that two offspring sampled in populations 1 and 2 share an HS relationship is expressed as follows: 

@@

which provides the same formulation described in Eq.\ref{HS-1}. Note that, from the third to forth lines, we assume that $\lambda_{l,1}$ and $\lambda_{l,2}$ ($l = 1, \ldots, M$) are independent variables (i.e., $\mathbb{E}[\lambda_{l,1}\lambda_{l,2}] = {\bar \lambda_{1}}{\bar \lambda_{2}}$), which implies variable reproductive potential of an identical parent before and after migration.

@@

In the derivation of $\pi_{\rm HS,bet}$ and $\pi_{\rm PO,bet}$ in {\bf Appendix 2}, we ignored covariation between migration and reproductive potential. Here, we consider the case where migrants have a distinguishable distribution of reproductive potential from nonmigrant parents. Let $\lambda_{i,1}$, $\lambda_{j,2}$, $\lambda_{M,l,1}$ and $\lambda_{M,l,2}$ be the reproductive potential of nonmigrant parents in population 1 and 2 and migrants in the populations 1 and 2 with mean ${\bar \lambda_{1}}$, ${\bar \lambda_{2}}$, ${\bar \lambda_{M,1}}$, and ${\bar \lambda_{M,2}}$, respectively. Under this setting, $\pi_{\rm HS,bet}$ can be expressed as follows:
\begin{align}
\pi_{\rm HS,bet} &\approx \frac{4}{M}  \frac{ \mathbb{E}\left[ \sum_{l=1}^M \lambda_{M,l,1} \sum_{l=1}^M \lambda_{M,l,2} \right]} {\mathbb{E}\left[ \left(\sum_{l=1}^{M} \lambda_{M,l,1} + \sum_{i=M+1}^{N_1} \lambda_{i,1}\right) \left(\sum_{l=1}^{M} \lambda_{M,l,2} + \sum_{j=M+1}^{N_2} \lambda_{j,2}\right) \right] }\nonumber\\ 
&= \frac{4}{M} \frac{ M^2 {\bar \lambda_{M,1}} {\bar \lambda_{M,2}} } { \left( M {\bar \lambda_{M,1}} + (N_1 - M) {\bar \lambda_1} \right) \left( M {\bar \lambda_{M,2}} + (N_2 - M) {\bar \lambda_2} \right)}.
\label{App3}
\end{align}
If ${\bar \lambda_{M,1}} = {\bar \lambda_{1}}$ and ${\bar \lambda_{M,2}} = {\bar \lambda_{2}}$, Eq.~\ref{App3} can be reduced Eq.\ref{App2}. This formulation includes reproductive potential terms such that the distinct reproductive potential between migrants and nonmigrants eliminates the usefulness of the proposed HS-based estimators (this is also applied to the proposed PO-based estimators).

>>>>

When both HS and PO pairs, which are found between populations, are available simultaneously, we can obtain efficient estimators by combined $\widehat{M_1}$ and $\widehat{M_2}$ for migration number: 
\begin{align}
\widehat{M_3} &= \frac{ N_{1} N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) },
\label{AP-1}
\end{align}
and by combining $\widehat{m_1}$ and $\widehat{m_2}$ for the migration rate: 
\begin{align}
\widehat{m_3} &= \frac{ N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right)}.
\label{AP-2}
\end{align}
For those estimators, $\widetilde{H}_{\rm HS,bet}$ and $\widetilde{H}_{\rm PO,bet}$ are weighted by $2n_{\rm O, 2}$ and $n_{\rm P, 2}$, respectively.

@@

We developed an individual-based model that tracks kinship relationships to evaluate the performance of the estimators. Here, the population structure was assumed to be identical to that in the development of the estimators. Population 1 and 2 comprised $N_1$ and $N_2$ parents with an equal sex ratio, and their offspring number was assumed to follow the geometric distribution with mean ${\bar \lambda_1}$ and ${\bar \lambda_2}$ (i.e., Poisson reproduction with mean $\lambda_{i,1}$ and $\lambda_{j,2}$, which follows the exponential distribution with mean ${\bar \lambda_1}$ and ${\bar \lambda_2}$), respectively. Migrant parents were selected randomly from population 1 at the end of the first year, and each offspring retained IDs of the parents, thereby making it possible to trace a HS and PO relationship.

@@

First, we evaluated the accuracy of ${\hat M}$s based on the relative bias. As expected, for most of the investigated parameter sets, we observed that their relative bias was less than 5\%, as shown in {\bf Table S1} in {\bf Supporting Information}. Thus, when the assumptions that (i) offspring and parents are randomly sampled, (ii) kinships are detected without any error, and (iii) the migration event does not depend on the degree of reproductive potential, are satisfied, it is reasonable to consider ${\hat M}$s nearly unbiased estimators.

@@

As the sample size or number of migrants increase, precision increases, and the shape of the distribution asymptotically becomes symmetric ({\bf Fig.~\ref{violins}}) because an increasing sample size or increasing number of migrants is likely to increase the observed number of kinship pairs found between the two populations ($\widetilde{H}_{\rm HS,bet}$ or $\widetilde{H}_{\rm PO,bet}$) and decreases the variance of those kinship pair numbers. In addition, the number of PO pairs in a given population ($\widetilde{H}_{\rm PO,1}$ or $\widetilde{H}_{\rm PO,2}$) contributes to the precision of estimating $N_1$ or $N_2$, respectively, thereby providing ${\hat M_4}$ and ${\hat M_5}$ with a relatively high precision.

Note that our simulation can handle case with and without invasive sampling. Invasive sampling potentially affects the level of $m$ because the sampled parents from population 1 have no chance to move to population 2; thus, such parents cannot be sampled. In this case, $m$ might be defined by $M/(N_1-n_{P,1})$ rather than $M/N_1$, although this does not affect the estimator ${\hat M}$s, as demonstrated in {\bf Table S1} ({\bf Supporting Information}).

@@

Our primary contributions are summarized as follows. First, we formulated the probabilities of kinship pairs randomly selected between two populations ($\pi_{\rm HS,bet}$ and $\pi_{\rm PO,bet}$), which provides migration number or rate estimators. While similar derivations may be found in the (nongenetic) MR method, these are limited to cases where sampling is non-invasive and adult individuals must be sampled. The proposed method can avoid these limitations, where such an advantages is characterized by the CKMR method for estimating population sizes. Second, we have demonstrated that the probabilities of kinship pairs are approximately independent of the reproductive potential; thus, information about it is not required to realize the estimation provided that migration is assumed to occur independently of the reproductive potential. This is a useful property of the estimators because it is natural that heterogeneity of reproductive potential within and between populations exits in iteroparous species but its observation is generally difficult. Third, we have demonstrated that estimators can be obtained using only genetic data (i.e., ${\hat M_5}$ and ${\hat m_4}$). There are several advantages in using the proposed estimator rather than estimating $\pi_{\rm HS,bet}$ (and/or $\pi_{\rm PO,bet}$) separately (via a genetic method) and $N_1$ (and/or $N_2$) (via a nongenetic method), including simplified sampling processes and analyzing designs and the availability of a unified framework of genetic analyses for detecting HS and PO pairs (similar discussion is found for estimating the ratio of effective breeding size to the census size, $N_{\rm b}/N$, in \cite{Akita_2020}).

@@

We believe there are several scenarios where the proposed estimator may be beneficial. The first case is where there is a large number of migrants between populations, e.g., $M>100$, which eliminates population differentiation and thus hampers the detection of the $M$ via population genetics method if $M$ is interpreted as an effective migration number. Information about movement between populations is essential to assess population dynamics in the conservation and management contexts, even if the migrant number is so large that it cannot be genetically assigned to two populations. The second case involves genetic monitoring conducted for each reproductive period (e.g., annually) because the proposed estimators consider information about contemporary migration that is explicitly specified the timing (e.g., year); thus, the time-series data for migration may reflect environmental changes, and this would provide insights into the underlying ecological processes. In addition, HS pairs found within the same cohort in the genetic population provide $N_{\rm b}$ \cite[]{wang2009new,WAPLES_2011}, which is also used to assess genetic health. The third is the case involves the development of integrated models that combine several data sources into a single analysis \cite[]{MAUNDER201361}. An example of this direction is to provide additional information about migration to ongoing CKMR projects for stock assessment in fisheries \cite[e.g.,][]{Bravington_2016,Hillary_2018,https://doi.org/10.1002/ece3.6296,ecolevol2021p,10.1093/icesjms/fsac002}, which would utilize kinship pairs for simultaneous estimation of population parameters. If the population structure is defined hypothetically, the data accumulated in such projects could be readily connected to our theory via likelihood methods. The fourth case is where sampling adults is difficult due to conservation practices or other reasons but instead sampling offspring is relatively easy. In such cases, although $N_2$ (and $N_1$) must be given externally to employ $\hat m_1$ (and $\hat M_1$), there is presently no existing method to estimate adult movement, especially when two populations cannot be genetically distinct, which indicates that the proposed method has the potential to expand the scope of current population monitoring techniques.

Finally, we discuss some caveats in terms of applying the proposed method. Our theory for developing the estimators assumes that kinships are detected without error. Many algorithms can be used to detect kinship pairs from single nucleotide polymorphisms or short tandem repeats \cite[e.g., ][]{Wang_2009Genetics, Huisman_2017}, although a HS pair requires many more DNA markers than a PO pair. In addition, iteroparous species may has several kinship types, e.g., half-uncle--nephew or half-cousins, which are expected to appear frequently and should be accurately differentiated from HS pairs. It is desirable to estimate in advance how many markers are required for kinship detection in the target populations associated with simulation of pedigree reconstruction (e.g., \cite{Anderson_2020}). The proposed estimators are limited to detecting parental movements on a period between breeding seasons in the given populations. Thus, the estimation of fine-scale spatiotemporal movements, which is available to integrate data sources \cite[]{Thorson_2021}, is beyond the scope of this paper. Although population dynamics models with coarser spatial resolution than the spatial scale of environmental layers are frequently used in assessment models, i.e., the target application of the proposed method, the proposed estimators require pre-specification of the population structure. Kinship relationships with sample location information potentially realize the ability to explore a plausible population structure by estimating the migration number or rate between hypothetical populations, which is expected to contribute the determination of a management unit. In addition, the current theory behind the estimators does not assume desynchronized reproduction within a population (e.g., skip spawning), correlation between individual mobility and fertility, and nonrandom sampling of parents (e.g., mothers with a large number of offspring are likely to be sampled). These issues will be the focus of future work.

@@

In the main text, we ignore the variation of reproductive potential among parents (i.e., both $\lambda_{i,1}$ and $\lambda_{j,2}$ are constant) to derive $\pi_{\rm HS,bet}$ and $\pi_{\rm PO,bet}$. Here, let $k_{i,1}$ and $k_{j,2}$ be the number of surviving offspring of parents $i$ and $j$ at sampling time in populations 1 and 2, and assumed to follow a kind of discrete distribution (e.g., Poisson or negative binomial distribution) with mean $\lambda_{i,1}$ and $\lambda_{j,2}$, respectively. Without loss of generality, we set the index such that parents with $i=1$ to $M$ in population 1 and $j=1$ to $M$ in population 2 are identical migrants. For example, the parent with $i=1$ reproduces $k_{1,1}$ offspring (in population 1) and then reproduce $k_{1,2}$ offspring (in population 2) after migration. Given $k_{i,1}$ and $k_{j,2}$, the conditional probability that two offspring sampled in populations 1 and 2 share a HS relationship is expressed as follows: 

@@

which provides the same formulation described in Equation~\ref{HS-1}. Note that, from the third to forth lines, we assume that $\lambda_{l,1}$ and $\lambda_{l,2}$ ($l = 1, \ldots, M$) are independent variables (i.e., $\mathbb{E}[\lambda_{l,1}\lambda_{l,2}] = {\bar \lambda_{1}}{\bar \lambda_{2}}$), which implies variable reproductive potential of an identical parent before and after migration.

@@

In the derivation of $\pi_{\rm HS,bet}$ and $\pi_{\rm PO,bet}$ in {\bf Appendix 2}, we ignored covariation between migration and reproductive potential. Here, we consider the case where migrants have a distinguishable distribution of reproductive potential from nonmigrant parents. Let $\lambda_{i,1}$, $\lambda_{j,2}$, $\lambda_{M,l,1}$ and $\lambda_{M,l,2}$ be the reproductive potential of nonmigrant parents in population 1 and 2 and migrants in the populations 1 and 2 with mean ${\bar \lambda_{1}}$, ${\bar \lambda_{2}}$, ${\bar \lambda_{M,1}}$, and ${\bar \lambda_{M,2}}$, respectively. Under this setting, $\pi_{\rm HS,bet}$ can be expressed as follows:
\begin{align}
\pi_{\rm HS,bet} &\approx \frac{4}{M}  \frac{ \mathbb{E}\left[ \sum_{l=1}^M \lambda_{M,l,1} \sum_{l=1}^M \lambda_{M,l,2} \right]} {\mathbb{E}\left[ \left(\sum_{l=1}^{M} \lambda_{M,l,1} + \sum_{i=M+1}^{N_1} \lambda_{i,1}\right) \left(\sum_{l=1}^{M} \lambda_{M,l,2} + \sum_{j=M+1}^{N_2} \lambda_{j,2}\right) \right] }\nonumber\\ 
&= \frac{4}{M} \frac{ M^2 {\bar \lambda_{M,1}} {\bar \lambda_{M,2}} } { \left( M {\bar \lambda_{M,1}} + (N_1 - M) {\bar \lambda_1} \right) \left( M {\bar \lambda_{M,2}} + (N_2 - M) {\bar \lambda_2} \right)}.
\label{App3}
\end{align}
If ${\bar \lambda_{M,1}} = {\bar \lambda_{1}}$ and ${\bar \lambda_{M,2}} = {\bar \lambda_{2}}$, Equation~\ref{App3} can be reduced Equation~\ref{App2}. This formulation includes reproductive potential terms such that the distinct reproductive potential between migrants and nonmigrants eliminates the usefulness of the proposed HS-based estimators (this is also applied to the proposed PO-based estimators).
